# Reproducible Research: Peer Assessment 1

## Loading and preprocessing the data
```{r}
   act <- read.csv("activity.csv")

   ok <- complete.cases(act)
  
   agrDate <- aggregate(act[ok,]$steps,by=list(act[ok,]$date), FUN = sum)

   names(agrDate)[1] <- "date"
   names(agrDate)[2] <- "steps"

   head(agrDate) 
```

## What is mean total number of steps taken per day?

### Histogram of total steps per day, excluding missing values
```{r fig.height=4, fig.width=12}
   hist(agrDate$steps,col="light green",main="Histogram of the total # steps thaen each day",
        xlab="Total # Steps / Day") 
```

### Mean of total steps taken per day
```{r}
   mean(agrDate$steps)
```

### Median of total steps taken per day
```{r}
   median(agrDate$steps)
```

## What is the average daily activity pattern?

### Aggregate steps by interval
```{r}
   agrInterval <- aggregate(act[ok,]$steps,by=list(act[ok,]$interval), FUN = mean)

   names(agrInterval)[1] <- "interval"
   names(agrInterval)[2] <- "steps"

   head(agrInterval)
```

### Time series plot of the 5-minute interval
```{r fig.height=4,fig.width=12}
 
   plot(agrInterval$steps,agrInterval$interval,type="l",main="Time series plot of the 5-minute interval",
        xlab="Intervals",ylab="Number of Steps",xlim=c(0, 220), ylim=c(0, 2500))
```

## Imputing missing values

### Total rows with missing values
```{r}
   nrow(act[!ok,])
```

### The missing values will be filled with the mean of respective interval
```{r}
   intervalNA <- unique(act[!ok,]$interval)

   ajusted <- act
   
   for(i in intervalNA)
   {
       newValue <- unique(agrInterval[agrInterval$interval == i,]$step)

       fillNA <- ajusted$interval == i & is.na(ajusted$step)

       ajusted[fillNA,1] <- newValue
   }
```

### Histogram of total steps per day, filling the missing values
```{r fig.height=4, fig.width=12}
   
   agrDateAjusted <- aggregate(ajusted$steps,by=list(ajusted$date), FUN = sum)

   names(agrDateAjusted)[1] <- "date"
   names(agrDateAjusted)[2] <- "steps"

   head(agrDateAjusted)

   hist(agrDateAjusted$steps,col="light blue",
        main="Histogram of the total # steps thaen each day, filling the missing values",
        xlab="Total # Steps / Day")
```

### Mean of total steps taken per day, filling the missing values
```{r}
   mean(agrDateAjusted$steps)
```

### Median of total steps taken per day, filling the missing values
```{r}
   median(agrDateAjusted$steps)
```

** Filling the missing values made the median igual the mean and the distribution will be a little bit less right skewed.** 

## Are there differences in activity patterns between weekdays and weekends?

### Create a new variable with the information of weekday and weekend
```{r }
   ajusted$period <- factor(NA,levels=c("weekday","weekend"))

   ajusted$period <- ifelse(weekdays(as.Date(ajusted$date),abbreviate=TRUE) %in% c("Sat","Sun"),
                            "weekend","weekday")   

   agrIntervalAjusted <- aggregate(ajusted$steps,by=list(ajusted$interval,ajusted$period), FUN = mean)

   names(agrIntervalAjusted)[1] <- "interval"
   names(agrIntervalAjusted)[2] <- "period"
   names(agrIntervalAjusted)[3] <- "steps"

   head(agrIntervalAjusted); tail(agrIntervalAjusted)
```

### Panel plot containing a time series plot with weekend and weekday data
```{r fig.height=10,fig.width=12}
   pwDay <- agrIntervalAjusted[agrIntervalAjusted$period == "weekday",]

   pwEnd <- agrIntervalAjusted[agrIntervalAjusted$period == "weekend",]

   par(mfrow=c(2,1))

   plot(pwEnd$steps,pwEnd$interval,type="l",main="weekend",xlab="Interval",
        ylab="Number of steps",
        xlim=c(0, 220), ylim=c(0, 2500))  

   plot(pwDay$steps,pwDay$interval,type="l",main="weekday",xlab="Interval",
        ylab="Number of steps",
        xlim=c(0, 220), ylim=c(0, 2500))
```
